Do Non-Cognitive Skills Affect Employment of Youth During the School-to-Work Transition?

### 2024-10-01 04_manuscript/paper_draft.qmd
[code] data prep and regressions
[code] new merge

---
title: "Do Non-Cognitive Skills Affect Employment Outcomes of Youth During the School-to-Work Transition?" 
author: "Garen Avanesian, Southern Federal University"
format: pdf
editor: visual
bibliography: references.bib
biblio-style: bwl-FU
citeproc: true
csl: "apa-with-abstract.csl"
---

@@

The study analyses the impact of non-cognitive skills on employment outcomes of youth aged 15-29 during the school-to-work transition in Russia. Based on the data of the Russian Longitudinal Monitoring Survey collected in 20XX. Adopting a mixed-effects model approach, the study aims to estimate if the effect of non-cognitive skills on employment is heterogeneous due to the income level and completed level of education. The study finds out that......

@@

@@

The transition from school to work represents a pivotal phase in the lives of young individuals, marked by decisions that significantly shape their future careers and overall lives. Importance of school-to-work transition is emphasized by the fact that it has long-term implications on socio-economic outcomes of individuals beyond youth period: while success in getting a first job that matches skills and expectations constitutes the further path to occupational success, early non-employment or poor transition to the world of work have long-lasting and negative implications on professional outcomes and beyond [@akkermans2021; @zacher2021; @baert2013; @luijkx2009; @verbruggen2015]. Moreover, the evidence points out that "association between having a high level of education and securing permanent full-time employment is weaker than expected" and level of education does not lead to the higher levels of job satisfaction [@chesters2020]. While there are many approaches to the definitions of school-to-work transition, their extensive survey is provided in [@blokker2023]. In this work, I rely on the one that is provided by ILO and defines it "as the passage of a young person (aged 15 to 29 years13) from the end of schooling to the first fixed-term or satisfactory employment" [@ilo2010, p. 4]. With that respect, each young person aged 15-29 is classified under the following categories: (1) transition not yet started for those who are either in school or not in school, inactive and with no intention to look for a job; (2) in transition, for those who are either unemployed, or in unsatisfactory self-employment temporary job, in unpaid family employment, or not in school but searching for a job; (3) transited, for those who are either in fixed-term employment or in satisfactory temporary or self-employment. While Sustainable Development Goals define youth as the population in the age 15 to 24 yeras old, ILO extended the upper bound to acknowledge that many young women and men remain in education till 24 years, and a wider age frame is required to capture employment experiences of youth post graduation.

@@

Current study explores the effect of non-cognitive skills on employment outcomes of youth aged 15-29 during school-to-work transition in Russia. The Russian data highlights that youth has the highest unemployment rates: in 2022, they accounted for 26% for 15-19 year olds, 13.6% for 20-24 year olds, and 5.2% for 25-29 year olds. This is compared to 2.8% of unemployment for individuals in the age of 40 to 44 year olds, the lowest value in the sample [@rosstat2024, p. 122]. Further, out of all higher education graduates between 2020-2022 who find a job, for 24% the obtained employment does not match the received profession. Finally, the available national data evidence that the labor market in the country demonstrates unprecedented spatial segregation: with the 4% as a country's average unemployment in 2022, it varies in the range between 1.6% in Yamalo-Nenets Autonomous Okrug and 30% in Republic of Ingushetia.[^1]

@@

@@

### Effect of Non-Cognitive Skills on Employment Outcomes

@@

## Conclusion

>>>>

---
title: "Do Non-Cognitive Skills Affect Employment of Youth During the School-to-Work Transition?" 
author: "Garen Avanesian, Southern Federal University"
format: 
  pdf:
    geometry: 
      - top=30mm
      - left=30mm
      - right=30mm
    documentclass: article
    number-sections: true
    colorlinks: true
    colorurl: true
editor: visual
bibliography: references.bib
biblio-style: bwl-FU
citeproc: true
csl: "apa-with-abstract.csl"
---

@@

The study analyses the impact of non-cognitive skills on employment of youth aged 15-29 during the school-to-work transition in Russia. Based on the data of the Russian Longitudinal Monitoring Survey collected in 20XX. Adopting a mixed-effects model approach, the study aims to estimate if the effect of non-cognitive skills on employment is heterogeneous due to the income level and completed level of education. The study finds out that......

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r wd set up, include=FALSE}
# set up the path
path = "/Users/karavan88/Documents/GitHub/youth_empl_ncs"

```

```{r source files, include=FALSE}
source(file.path(path, "user_profile.R"))
source(file.path(rCodes, "0200_load_packages.R"))
# source(file.path(rCodes, "02_descriptive.R"))
# source(file.path(rCodes, "03_model.R"))

```

@@

The transition from school to work represents a pivotal phase in the lives of young individuals, marked by decisions that significantly shape their future careers and overall lives. Importance of school-to-work transition is emphasized by the fact that it has long-term implications on socio-economic outcomes of individuals beyond youth period: while success in getting a first job that matches skills and expectations constitutes the further path to occupational success, early non-employment or poor transition to the world of work have long-lasting and negative implications on professional outcomes and beyond [@akkermans2021; @zacher2021; @baert2013; @luijkx2009; @verbruggen2015]. Moreover, the evidence points out that "association between having a high level of education and securing permanent full-time employment is weaker than expected" and level of education does not lead to the higher levels of job satisfaction [@chesters2020]. While there are many approaches to the definitions of school-to-work transition, their extensive survey is provided in [@blokker2023]. In this work, I rely on the one that is provided by ILO and defines it "as the passage of a young person (aged 15 to 29 years) from the end of schooling to the first fixed-term or satisfactory employment" [@ilo2010, p. 4]. With that respect, each young person aged 15-29 is classified under the following categories: (1) transition not yet started for those who are either in school or not in school, inactive and with no intention to look for a job; (2) in transition, for those who are either unemployed, or in unsatisfactory self-employment temporary job, in unpaid family employment, or not in school but searching for a job; (3) transited, for those who are either in fixed-term employment or in satisfactory temporary or self-employment. While Sustainable Development Goals define youth as the population in the age 15 to 24 yeras old, ILO extended the upper bound to acknowledge that many young women and men remain in education till 24 years, and a wider age frame is required to capture employment experiences of youth post graduation.

@@

Current study explores the effect of non-cognitive skills on employment outcomes of youth aged 15-29 during school-to-work transition in Russia. The Russian official statistics highlights that youth has the highest unemployment rates: in 2022, they accounted for 26% for 15-19 year olds, 13.6% for 20-24 year olds, and 5.2% for 25-29 year olds. This is compared to 2.8% of unemployment for individuals in the age of 40 to 44 year olds, the lowest value by the age cohort breakdown [@rosstat2024, p. 122]. Further, out of all higher education graduates between 2020-2022 who found a job, for 24% the obtained employment does not match the received profession. Finally, the available national data evidence that the labor market in the country demonstrates unprecedented spatial segregation: with the 4% as a country's average unemployment in 2022, it varies in the range between 1.6% in Yamalo-Nenets Autonomous Okrug and 30% in Republic of Ingushetia.[^1]

@@

```{r}

```

@@

### Effect of Non-Cognitive Skills on Employment

@@

## Conclusions
